Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-03 paper/graph.tex
More edits
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@599 3fc257bb-8550-0410-81b3-a5c103da83f7
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> 1 Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> 1 Aggregate $E_{new}$. \\
\> 1 Add $E_{new}$ to $E_{old}$. \\
\> 1 Convert $E_{old}$ from KV to KMV. \\
\> 1 Reduce $E_{old}$: \> Remove duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

Input:  \\
\> Source vertex $v_s$.\\
\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
$U$ = new MapReduce object to store updates to distances. \\
Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
\> Input:  $v_s$ \\
\> Output:  Key = $v_s$; Value = 0. \\
while not done: \\
\> done = 1; \\
\> Aggregate $U$. \\
\> Add $U$ to $V$. Empty $U$.\\
\> Convert $V$:  Key is vertex $v_i$. \\
\> Reduce $V$:  Pick best candidate distance for each vertex. \\
\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
\> if not done \\
\> \> Add $U$ to $E$.  Empty $U$. \\
\> \> Convert $E$. \\
\> \> Reduce $E$:  Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
\> \> \> \> and distance $d_i$ if updated in previous step. \\
\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

@@

>>>>

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> 1 Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> 1 Aggregate $E_{new}$. \\
\> 1 Add $E_{new}$ to $E_{old}$. \\
\> 1 Compress $E_{old}$: \> Remove duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

Input:  \\
\> Source vertex $v_s$.\\
\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
$U$ = new MapReduce object to store updates to distances. \\
Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
\> Input:  $v_s$ \\
\> Output:  Key = $v_s$; Value = 0. \\
while not done: \\
\> done = 1; \\
\> Aggregate $U$. \\
\> Add $U$ to $V$. Empty $U$.\\
\> Compress $V$:  Pick best candidate distance for each vertex. \\
\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
\> if not done \\
\> \> Add $U$ to $E$.  Empty $U$. \\
\> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
\> \> \> \> and distance $d_i$ if updated in previous step. \\
\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

@@

 \caption{Improved MapReduce algorithm for single-source shortest path (SSSP)
 determination.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}

### 2010-05-03 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@499 3fc257bb-8550-0410-81b3-a5c103da83f7
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

MapReduce is the programming paradigm popularized by Google
researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
enable rapid development and deployment of analysis programs to
operate on massive data sets residing on Google's large distributed
clusters.  They introduced a novel way of thinking about certain kinds
of large-scale computations as a "map" operation followed by a
"reduce" operation.  The power of the paradigm is that when cast in
this way, a nominally serial algorithm now becomes two highly parallel
operations working on data local to each processor, sandwiched around
an intermediate data-shuffling operation that requires inter-processor
communication.  The user need only write serial code for the
application-specific map and reduce functions; the parallel data
shuffle can be encapsulated in a library since its operation is
independent of the application.

@@

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data \cite{Tu}, graph
algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
The paper by Tu et al \cite{Tu} was particularly insightful to us,
because it described how MapReduce could be implemented on top of the
ubiquitous distributed-memory message-passing interface (MPI), and how
the intermediate data-shuffle operation is conceptually identical to
the familiar $MPI_Alltoall$ operation.  Their implementation of
MapReduce was within a Python wrapper to simplify the writing of user
programs.  The paper motivated us to develop our own C++ library built
on top of MPI for use in graph analytics, which we initially released
as open-source software in mid-2009 \cite{MRMPI}.  We have since
worked to optimize several of the library's underlying algorithms and
to enable its operation in out-of-core mode on larger data sets.
These algorithmic improvements are described in this paper and are
part of the current downloadable version \cite{MRMPI}.

@@

\item {\it C++ library using MPI for inter-processor communication:}
The user writes a (typically) simple main program which runs on each
processor of a parallel machine, making calls to the MR-MPI library.
For map and reduce operations, the library calls back to user-provided
map() and reduce() functions.  The use of C++ allows precise control
over the memory and format of data allocated by each processor during
a MapReduce.  Library calls for performing a map, reduce, or data
shuffle, are synncrhonous, meaning all the processors participate and
finish the operation before proceeding.  Similarly, the use of MPI
within the library is the traditional mode of $MPI_Send()$ and
$MPI_Recv$ calls between processor pairs using large aggregated
messages to improve bandwidth performance and reduce latency costs.  A
recent paper by \cite{Dongarra} also outlines the MapReduce formalism
from an MPI perspective, though they advocate a more asyncrhonous
approach, using one-way communication of small messages.

\item {Small, portable:} The entire MR-MPI library is a few thousand
lines of standard C++ code.  For parallel operation, the program is
linked with MPI, a standard message passing library available on all
distributed memory machines.  For serial operation, a dummy MPI
library (provided) can be substituted.

\item {\it In-core or out-of-core operation:} Each MapReduce object a
processor defines, allocates per-processor "pages" of memory, where
the page size is determined by the user.  Typical MapReduce operations
can be performed using a few such pages.  If the data set fits in a
single page (per processor), then the library performs its operations
in-core.  If the data set exceeds the page size, then processors each
write to temporary disk files (to local disk or a parallel file
system) as needed and subsequently read from them.  This allows
processing of data sets larger than the aggregate memory of all the
processors, i.e. up to the available aggregate disk space.

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
$MPI_Allreduce$ are often useful in determining the status of an
iterative graph algorithm, as described in Section \ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed back to all callback functions, so it is easy for the user
program to store ``state'' on each processor, accessible during the
map and reduce operaions.  For example, various flags can be stored
that alter the operation of a map or reduce function, as can richer
data structures, that accumulate the results.

\item {\it C++, C, and Python interfaces:} A C++ interface to the
MR-MPI library means a user program instantiates and then invokes
methods in one or more MapReduce objects.  A C interface means the
library can also be called from C or other hi-level languages such as
Fortran.  A C interface also means the library can be easily wrapped
by Python via the Python ``ctypes'' module.  The library can then be
called from a Python script, allowing the user to write map() and
reduce() callback functions in Python.  If a machine supports running
Python in parallel, a parallel MapReduce can also be run in this mode.

@@

The remainder of the paper is organized as follows.  The next two
sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section \ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section \ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
ranging in size from 1 million to 1 trillion vertices or edges.  In
this section, we highlight the performance and complexity trade-offs
of a MapReduce approach versus other more special-purpose algorithms.
The latter generally perform better but are harder to implement
efficiently on distributed memory machines, due to the required
explicit management of parallelism, particularly for large out-of-core
data sets.  Section \ref{sec:lessons} summarizes some lessons learned
from the implementation and use of our library.

>>>>

MapReduce is the programming paradigm popularized by Google
researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
enable rapid development and deployment of analysis programs to
operate on massive data sets residing on Google's large distributed
clusters.  They introduced a novel way of thinking about certain kinds
of large-scale computations as a ``map'' operation followed by a
``reduce'' operation.  The power of the paradigm is that when cast in
this way, a nominally serial algorithm now becomes two highly parallel
operations working on data local to each processor, sandwiched around
an intermediate data-shuffling operation that requires inter-processor
communication.  The user need only write serial code for the
application-specific map and reduce functions; the parallel data
shuffle can be encapsulated in a library since its operation is
independent of the application.

@@

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data \cite{Tu}, graph
algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
The paper by Tu et al \cite{Tu} was particularly insightful to us,
because it described how MapReduce could be implemented on top of the
ubiquitous distributed-memory message-passing interface (MPI), and how
the intermediate data-shuffle operation is conceptually identical to
the familiar MPI\_Alltoall operation.  Their implementation of
MapReduce was within a Python wrapper to simplify the writing of user
programs.  The paper motivated us to develop our own C++ library built
on top of MPI for use in graph analytics, which we initially released
as open-source software in mid-2009 \cite{MRMPI}.  We have since
worked to optimize several of the library's underlying algorithms and
to enable its operation in out-of-core mode on larger data sets.
These algorithmic improvements are described in this paper and are
part of the current downloadable version \cite{MRMPI}.

@@

\item {\it C++ library using MPI for inter-processor communication:}
The user writes a (typically) simple main program which runs on each
processor of a parallel machine, making calls to the MR-MPI library.
For map and reduce operations, the library calls back to user-provided
map() and reduce() functions.  The use of C++ allows precise control
over the memory and format of data allocated by each processor during
a MapReduce.  Library calls for performing a map, reduce, or data
shuffle, are synchronous, meaning all the processors participate and
finish the operation before proceeding.  Similarly, the use of MPI
within the library is the traditional mode of MPI\_Send and
MPI\_Recv calls between processor pairs using large aggregated
messages to improve bandwidth performance and reduce latency costs.  A
recent paper by \cite{Dongarra} also outlines the MapReduce formalism
from an MPI perspective, though they advocate a more asyncrhonous
approach, using one-way communication of small messages.

\item {Small, portable:} The entire MR-MPI library is a few thousand
lines of standard C++ code.  For parallel operation, the program is
linked with MPI, a standard message passing library available on all
distributed memory machines.  For serial operation, a dummy MPI
library (provided) can be substituted.  As a library, it can be
embedded in other codes \cite{Titan,TitanURL} to enable them to
perform MapReduces.

\item {\it In-core or out-of-core operation:} Each MapReduce object that a
processor defines allocates per-processor ``pages'' of memory, where
the page size is determined by the user.  Typical MapReduce operations
can be performed using a few such pages.  If the data set fits in a
single page (per processor), then the library performs its operations
in-core.  If the data set exceeds the page size, then processors each
write to temporary disk files (to local disk or a parallel file
system) as needed and subsequently read from them.  This allows
processing of data sets larger than the aggregate memory of all the
processors, i.e. up to the available aggregate disk space.

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section \ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed back to all callback functions, so it is easy for the user
program to store ``state'' on each processor, accessible during the
map and reduce operations.  For example, various flags can be stored
that alter the operation of a map or reduce function, as can richer
data structures, that accumulate the results.

\item {\it C++, C, and Python interfaces:} A C++ interface to the
MR-MPI library means a user program instantiates and then invokes
methods in one or more MapReduce objects.  A C interface means the
library can also be called from C or other high-level languages such as
Fortran.  A C interface also means the library can be easily wrapped
by Python via the Python ``ctypes'' module.  The library can then be
called from a Python script, allowing the user to write map() and
reduce() callback functions in Python.  If a machine supports running
Python in parallel, a parallel MapReduce can also be run in this mode.

@@

The remainder of the paper is organized as follows.  The next two
sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section \ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section \ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
%KDDKDD CHECK SIZES
ranging in size from 1 million to 1 trillion vertices or edges.  In
this section, we highlight the performance and complexity trade-offs
of a MapReduce approach versus other more special-purpose algorithms.
The latter generally perform better but are harder to implement
efficiently on distributed memory machines, due to the required
explicit management of parallelism, particularly for large out-of-core
data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
from the implementation and use of our library.

### 2010-05-03 paper/mr.tex
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

A typical MR-MPI program makes at least 3 calls to the MR-MPI library,
to perform a {\it map()}, {\it collate()}, and {\it reduce()}
operation.  In a {\it map()}, zero or more key/value pairs are
generated by each processor.  Typically, this is done using data read
in from files, but a {\it map()} may generate data itself or process
existing KV pairs to create new ones.  The KV pairs produced are
stored locally by each processor; a {\it map()} thus requires no
inter-processor communication.  Users call the library with a count of
tasks to perform and a pointer to a user function; the MR-MPI {\it
map()} operation invokes the user function multiple times as a
callback.  Depending on which variant of {\it map()} is called, the
user function may be passed a file name, a chunk of bytes from a large
file, or a task ID.  Options for assigning map tasks to processors.

The {\it collate()} operation (or data shuffle in Hadoop) identifies
unique keys and collects all the values associated with those keys to
create KMV pairs.  This is done in two stages, the first of which
requires communication, since KV pairs with the same key may be owned
by any processor.  Each processor hashes each of its keys to determine
the processor who ``owns'' it.  The $N$-byte length key is hashed into
a 32-bit value whose remainder modulo $P$ generates the owning
processor ID.  Alternatively, the user can provide a hash function
which converts the key into a processor ID.  Each processor then sends
each of its KV pairs to the owning processor.

After receiving new KV pairs, the second stage is an on-processor
computation, requiring no further communication.  Each processor
reorganizes its KV pairs into KMV pairs, one for each unique key it
owns.  This is done using a hash table, rather than a sort.  A sort
requires $\log(N)$ passes through the $N$ KV pairs.  the list of KMV
pairs can be created in 2 passes through the KV pairs, one to populate
the hash table with needed count and length information, and the
second pass to copy the key and value datums into the appropriate
location in a new KMV data structure.  Since the lookup of a key in a
well-formed hash table is a constant-time $O(1)$ operation, the cost
of the data reorganization is $O(N)$, rather than $N\log(N)$ for full
sort.

@@

Several related MapReduce operations are provided by the library.  For
example, the {\it compress()} operation combines on-processor KV pairs
to eliminate duplicate keys, forming new KV pairs with an aggregated
value.  The { \it convert()} operation turns a list of KV pairs into
KMV pairs, with one value per key.  The {collapse()} operation turns
$N$ KV pairs into one KMV pair, with the keys and values of the KV
pairs becoming $2N$ values of a single multivalue assigned to a new
key on the KMV.  The {\it gather()} operation collects KV pairs from
all processors to a subset of processors; it is useful for doing
output from a single or subset of processors.  Library calls for
sorting datums by key or value or within multivalues are also
provided.  These routines invoke the C-library quicksort() function to
compute the sorted ordering of the KV pairs (or values within a
multivalue), using a user-provided comparison function.  The KV pairs
(or values in a multivalue) are then copied one-by-one into a new data
structure in sorted order.

The purpose of providing an interface to various lo-level operations,
is that a user program can string them together in various ways to
produce interesting MapReduce algorithms.  For example, output from a
{\it reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  KV pairs from mutiple MapReduce objects can be combined
together to perform new sequences of {\it map()}, {\it collate()}, and
{\it reduce()} operations.

The above discussion assumed that the KV or KMV pairs stored by a
processor fit in its physical memory.  If this is the case, then we
refer to this as ``in-core'' processing and no disk files are written
or read by any of the processors, aside from initial input data if it
exists or final output data if it is generated.  Note that the
aggregate physical memory of large parallel machines can be multiple
Tbytes, which allows for large data sets to be processed in-core,
assuming the KV and KMV pairs remain evenly distributed across
processors throughout the sequence of MapReduce operations.  The use
of randomized hashing to assign keys to processors is designed for
such load-balancing.  What happens when data sets do not
fit in available memory, is the subject of the next section.

>>>>

A typical MR-MPI program makes at least three calls to the MR-MPI library,
to perform a {\it map()}, {\it collate()}, and {\it reduce()}
operation.  In a {\it map()}, zero or more key/value pairs are
generated by each processor.  Typically, this is done using data read
in from files, but a {\it map()} may generate data itself or process
existing KV pairs to create new ones.  The KV pairs produced are
stored locally by each processor; a {\it map()} thus requires no
inter-processor communication.  Users call the library with a count of
tasks to perform and a pointer to a user function; the MR-MPI {\it
map()} operation invokes the user function multiple times as a
callback.  Depending on which variant of {\it map()} is called, the
user function may be passed a file name, a chunk of bytes from a large
file, a task ID, or a KV pair.  Options for assigning map tasks to processors
are specified by the user and include assigning consecutive 
chunks of tasks to processors, striding the tasks across processors,
or using a master-slave model that is useful when tasks have widely
varying workloads.

The {\it collate()} operation (or data shuffle in Hadoop) identifies
unique keys and collects all the values associated with those keys to
create KMV pairs.  This is done in two stages, the first of which
requires communication, since KV pairs with the same key may be owned
by any processor.  Each processor hashes each of its keys to determine
which processor will ``own'' it.  The $k$-byte length key is hashed into
a 32-bit value whose remainder modulo $P$ processors generates the owning
processor ID.  Alternatively, the user can provide a hash function
which converts the key into a processor ID.  Each processor then sends
each of its KV pairs to the owning processor.

After receiving new KV pairs, the second stage is an on-processor
computation, requiring no further communication.  Each processor
reorganizes its KV pairs into KMV pairs, one for each unique key it
owns.  This is done using a hash table, rather than a sort.  A sort
requires $\log(N)$ passes through the $N$ KV pairs.  With hashing,
the list of KMV
pairs can be created in two passes through the KV pairs, one to populate
the hash table with needed count and length information, and the
second pass to copy the key and value datums into the appropriate
location in a new KMV data structure.  Since the lookup of a key in a
well-formed hash table is a constant-time $O(1)$ operation, the cost
of the data reorganization is $O(N)$, rather than $O(N\log(N))$ for full
sorting.  It also has the added benefit that the number of values in each
KMV pair is known and is passed to the user function during a
reduce.  For some reduce operations, this count is all the information a
reduce requires; thus the values need not be looped over.  This feature is not
available in Hadoop-style data reorganizations, which sort the values; 
with sorting, the number of values associated with the same key is not known 
without interating over the values.

@@

Several related MapReduce operations are provided by the library and
exposed in the library interface.  For example, the {\it collate()} function
described above
calls two other functions:  {\it aggregate()}, which performs the all-to-all
communication, and {\it convert()}, which turns a list of KV pairs into
KMV pairs.  
Both {\it aggregate()} and {\it convert()} are exposed in the MR-MPI
interface for direct access by the user.
The {\it compress()} function allows on-processor operations to be 
performed; it is equivalent to a {\it convert()} with on-processor KVs as
input, followed by a {\it reduce()}.  
The {\it clone()} operation turns a list of KV pairs into
KMV pairs, with one value per key.  The {\it collapse()} operation turns
$N$ KV pairs into one KMV pair, with the keys and values of the KV
pairs becoming $2N$ values of a single multivalue assigned to a new
key on the KMV.  The {\it gather()} operation collects KV pairs from
all processors to a subset of processors; it is useful for doing
output from a single or subset of processors.  Library calls for
sorting datums by key or value or within multivalues are also
provided.  These routines invoke the C-library {\it quicksort()} function to
compute the sorted ordering of the KV pairs (or values within a
multivalue), using a user-provided comparison function.  The KV pairs
(or values in a multivalue) are then copied one-by-one into a new data
structure in sorted order.

The interface to various low-level operations is provided so
that a user's program can string them together in various ways to
produce interesting MapReduce algorithms.  For example, output from a
{\it reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  KV pairs from mutiple MapReduce objects can be combined
together to perform new sequences of {\it map()}, {\it collate()}, and
{\it reduce()} operations.

The above discussion assumed that the KV or KMV pairs stored by a
processor fit in its physical memory.  If this is the case, then we
refer to this as ``in-core'' processing and no disk files are written
or read by any of the processors, aside from initial input data if it
exists or final output data if it is generated.  Note that the
aggregate physical memory of large parallel machines can be multiple
terabytes, which allows for large data sets to be processed in-core,
assuming the KV and KMV pairs remain evenly distributed across
processors throughout the sequence of MapReduce operations.  The use
of randomized hashing to assign keys to processors is designed for
such load-balancing.  In the next section, we discuss what happens 
when data sets do not fit in available memory.
